### 2014-11-11 locks.tex
Update Related Work.

\begin{topfig}
\footnotesize\begin{tabular}{ | l || c | c | c | c | c |}
\hline
\multirow{2}{*}{\textbf{Program}} & \multicolumn{5}{c |}{\textbf{Queue
   Instructions}} \\ \cline{2-6}
& 1 & 2 & 4 & 8 & 16 \\ \hline
\hline
SSSP - Regular & 298K & 305K & 309K & 286K & 325K \\ \hline
SSSP - Coordinated & 961K & 919K & 903K & 900K & 922K \\ \hline
SSSP - Buffered & 939K & 961K & 999K & 926K & 1M \\ \hline
\hline
HT - Buffered & 38M & 49M & 50M & 52M & 54M \\ \hline
HT - Coordinated & 57M & 63M & 95M & 89M & 82M \\ \hline
HT - Local-Only & 57M & 63M & 91M & 80M & 74M \\ \hline
\hline
LBP - Regular & 23M & 26M & 24M & 17M & 12M \\ \hline
SBP - Coordinated & 95M & 93M & 88M & 81M & 75M \\ \hline
\hline
MiniMax - Regular & 24M & 26M & 27M & 26M & 27M \\ \hline
MiniMax - Coordinated & 51M & 51M & 52M & 51M & 52M \\ \hline
\end{tabular}
\scap{queue_instructions}{Number of queue \scare{instructions} per program. An
   operation to a regular queue counts as 1 instruction, while an operation to
   a priority queue costs the number of
   \emph{percolate-up}/\emph{percolate-down} steps performed.}
\end{topfig}

>>>>

\begin{topfig}
\footnotesize\begin{tabular}{ | l || c | c | c | c | c |}
\hline
\multirow{2}{*}{\textbf{Program}} & \multicolumn{5}{c |}{\textbf{Queue
   Instructions}} \\ \cline{2-6}
& 1 & 2 & 4 & 8 & 16 \\ \hline
\hline
SSSP - Regular & 298K & 300K & 313K & 284K & 317K \\ \hline
SSSP - Coordinated & 953K & 960K & 861K & 911K & 987K \\ \hline
SSSP - Buffered & 951K & 1M & 1M & 1M & 1M \\ \hline
\hline
HT - Buffered & 38M & 49M & 52M & 74M & 55M \\ \hline
HT - Coordinated & 57M & 64M & 95M & 87M & 83M \\ \hline
HT - Local-Only & 57M & 63M & 90M & 85M & 77M \\ \hline
\hline
LBP - Regular & 23M & 26M & 24M & 17M & 12M \\ \hline
SBP - Coordinated & 96M & 92M & 90M & 78M & 74M \\ \hline
\hline
MiniMax - Regular & 24M & 25M & 26M & 26M & 27M \\ \hline
MiniMax - Coordinated & 51M & 52M & 52M & 52M & 52M \\ \hline
\hline
N Queens - Regular & 12K & 24K & 25K & 365K & 5M \\ \hline
N Queens - Coordinated & 57M & 11M & 904K & 145K & 44K \\ \hline
\end{tabular}
\scap{queue_instructions}{Number of queue \scare{instructions} per program. An
   operation to a regular queue counts as 1 instruction, while an operation to
   a priority queue costs the number of
   \emph{percolate-up}/\emph{percolate-down} steps performed.}
\end{topfig}

### 2014-11-11 related.tex
More updates.
Update Related Work.

There has been an increasing interest in declarative 
languages. MapReduce~\cite{Dean:2008:MSD:1327452.1327492}, for instance, is a
popular data-centric programming model that is optimized for large clusters.
Intrinsic to its popularity is the simplicity of its scheduling and data sharing model.
In order to facilitate the writing of programs over large
datasets, SQL-like languages such as
PigLatin~\cite{Olston:2008:PLN:1376616.1376726} have been developed.
PigLatin builds on top of MapReduce and allows the programmer to write complex
data-flow graphs, raising the abstraction and easy of programmability of
MapReduce programs.
An alternative to PigLatin/MapReduce is
Dryad~\cite{Isard:2007:DDD:1272996.1273005}
that allows programmers to design arbitrary computation patterns using the DAG
abstraction. It combines computational vertices with communication channels
(edges). Dryad programs are automatically scheduled to run on multiple
computers/cores and data is partitioned automatically during runtime.

Many programming languages follow the so-called \emph{coordination
paradigm}~\cite{Papadopoulos98coordinationmodels}, a form of distributed
programming that divides execution in two parts: \emph{computation}, where the actual
computation is performed, and \emph{coordination}, which deals with
communication and cooperation between processing units. This paradigm attempts
to clearly distinguish between these two parts by providing abstractions for
coordination in an attempt to provide architecture and system-independent forms
of communication.

@@

Linda and Delirium are limited in the sense that the programmer can only
coordinate the scheduling of processing units, while the placement of data is
left to the implementation. But even more importantly, CLM differs from those
languages because coordination acts on data instead of processing units.
The abstraction is then raised by considering data and algorithmic aspects of
the program instead of focusing how processing units are used.
Furthermore, the LM language is both a coordination language and a computation
language and there is no distinction between the two components.

>>>>

\paragraph{Declarative Languages}
Declarative programming paradigms such as \emph{logic} and \emph{functional}
programming have been extensively exploited for their implicit parallelism.
In logic languages such as Prolog, researchers took advantage of the
non-determinism of proof-search to evaluate subgoals in parallel with models
such as \emph{or-parallelism}~\cite{ali-86} and
\emph{and-parallelism}~\cite{Shen-92}.  In functional languages, the stateless
nature of computation allows multiple expressions to evaluate safely in
parallel. This has been explored in several languages such as
NESL~\cite{Blelloch:1996:PPA:227234.227246}, Id~\cite{Nikhil93anoverview}, and
more recently Data Parallel Haskell~\cite{nepal}.

\paragraph{Data Centric Languages}
Recently, there has been an increasing interest in declarative data-centric
languages.  MapReduce~\cite{Dean:2008:MSD:1327452.1327492}, for instance, is a
popular data-centric programming model that is optimized for large clusters.
Intrinsic to its popularity is the simplicity of its scheduling and data sharing
model.  In order to facilitate the writing of programs over large datasets,
SQL-like languages such as PigLatin~\cite{Olston:2008:PLN:1376616.1376726} have
been developed.  PigLatin builds on top of MapReduce and allows the programmer
to write complex data-flow graphs, raising the abstraction and easy of
programmability of MapReduce programs.  An alternative to PigLatin/MapReduce is
Dryad~\cite{Isard:2007:DDD:1272996.1273005} that allows programmers to design
arbitrary computation patterns using the DAG abstraction. It combines
computational vertices with communication channels (edges) that are
automatically scheduled to run on multiple computers/cores.

GraphLab~\cite{GraphLab2010} is a C++ framework for developing parallel machine
learning algorithms. GraphLab allows nodes to have read/write access to
different scopes through different concurrent access models in order to balance
performance and data consistency.  GraphLab also provides different schedulers
that dictate the order in which node's are computed, allowing the programmer to
optimize the program. Later in this paper, we will show how one GraphLab
scheduler can be implemented in CLM through the use of coordination facts.

\paragraph{Coordination Languages}
Many programming languages follow the so-called \emph{coordination
paradigm}~\cite{Papadopoulos98coordinationmodels}, a form of distributed
programming that divides execution in two parts: \emph{computation}, where the actual
computation is performed, and \emph{coordination}, which deals with
communication and cooperation between processing units. This paradigm attempts
to clearly distinguish between these two parts by providing abstractions for
coordination in an attempt to provide architecture and system-independent forms
of communication.

@@

Linda and Delirium are limited in the sense that the programmer can only
coordinate the scheduling of processing units, while the placement of data is
left to the implementation. But even more importantly, CLM differs from those
languages because coordination acts on data instead of processing units.
The abstraction is then raised by considering data and algorithmic aspects of
the program instead of focusing how processing units are used.
Furthermore, the CLM language is both a coordination language and a computation
language and there is no distinction between the two components.

\paragraph{Forward-Chaining Logic Programming}
